---
title: "SALEM_2_edited"
author: "Alex Salem"
date: "May 12, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Preliminary Information:

###Loading libraries:
```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(gapminder)
library(skimr)
library(scales)
```

###EDA:
```{r}
gapminder
glimpse(gapminder)
```


###Original Graph:
```{r task number 2 part 2}
library(ggbeeswarm)

ggplot(gapminder, aes(x = continent, y = gdpPercap)) +
  geom_quasirandom(color = "lightseagreen", alpha = .5) +
  scale_y_log10() + #from data camp 
  labs(x = "Continent", y = "GDP per capita") +
  ggtitle("Spread of GDP per capita across continent")
    
```

###Original criticism:
I think this graph needs a better y-axis. The log scale needs to be better explained and needs a better label and scale.

###New Graph:
```{r}
#filter year:
gapminder_2007 <- gapminder %>% 
  filter(year == 2007)

ggplot(gapminder_2007, aes(x = continent, y = gdpPercap)) +
  geom_quasirandom(color = "lightseagreen", alpha = .6) +
  scale_y_log10(labels = comma) + #from data camp 
  labs(x = "Continent", y = "GDP per capita", caption = "Note: We are using a log scale for the y-axis. \nThus, each labelled tick mark on y-axis is the previous tick mark multiplied by ten.") +
  ggtitle("Spread of GDP per capita across countries \nwithin continents, in 2007")
```

##Changes made:
* Limited the data to looking only at one year: 2007. This makes more sense than including all the years at once.
* Changed the y-axis scale from using scientific notation to using standard comma separated numbers so that it's easier to read.
* Added a caption explaining the log scale, so that an unfamiliar audience will understand it.
* Increased the alpha value for the points.

##Graph Brief:

###Type of Graph:
This is a *beeswarm plot*.

###Data:
The data comes from the gapminder data set. This data set has 6 variables: country, continent, year, lifeExp (life expectation), population, and gdpPercap (GDP per capita). There are 1704 total observations.

###Audience:
Here, I am aiming for a fairly general audience--anyone with familiarity with basic plots and how to read them should be able to process this, but no further previous information on this specific data is required. With the previous graph, I think the reader needed to understand what a log scale meant as well, which is why I added the explanation of that to my new graph.

###Representation Description: 
In general, a beeswarm represents each observation within a cateogry (along the x-axis) as a point. The point's vertical height (its location on the y-axis), is some value of a continous variable. It is used to show the spread of observations within a category.

###How to read it & What to look for:
A newbie should first identify the title and x- and y-axis labels. This should give a sense of what category is being considered, and what variable of that cateogry is being assessed. Then, the viewer should check out a few specific points--identify their x- and y-values to gauge the range of observations. Then, the viewer should look more generally at the shape of the beeswarm(s). What trend do you see in each cateogry? What is the spread of observations in that cateogory like? How does this compare to other values in that cateogory (if there is more than one)?

Consider this graph. It shows the spread of GDP per capita across countries within continents, in 2007. Thus in this graph, each observation (point) represents the GDP per capita for one country within the specified continent along the x-axis. Consider the beeswarm for Africa. It looks like a number of observations are clustered around 1000 GDP per capita, and another smaller cluster is around 10,000 GDP per capita. 

###Presentation tips:
**Annotation**: Annotation can be used to indicate certain important points in the beeswarm, or to show an overall comparison value. Here, we didn't make use of annotation.

**Color**: Color can be used to layer on another variable to the beeswarm. We didn't make that use of color here, but did color all the points one color to make it prettier and easier to read. Additionally, since there are many data points in each beeswarm, I used `alpha = .6` to make the color semiopaque, so that overlap of points was more clear.

**Arrangement**: Here, we used a log scale for the y-axis, since this data had such a large spread. This allows us to see the points more clearly. Without it they are "bunched up". You can also vary the arrangement of points using the "smiley" method discussed below.

###Variations and alternatives:
There is one main variation I'm aware of which is adding the "smiley method". You can add that with `geom_quasirandom(method = "smiley")`. This creates sort of "rows" of dots in the beeswarm, which can make the presentation neater. 

###How I created it:
I used ggplot2 in R, and the ggbeeswarm package. I created the x- and y- variables using `aes(x = continent, y = gdpPercap)`. Then, I was able to create a ggplot with a beeswarm using the `geom_quasirandom()` layer. Then I layered on the log scale for the y-axis with `scale_y_log10(labels = comma)` (the `labels = comma` part wrote out the y-values fully instead of using scientific notation). Finally, I added the x- and y-axis labels and the title using `labs()` and `ggtitle()` respectively. 